---
title: "Homework 2"
author: Yuanxin Zhang
output: github_document
---

This is my solution to Homework 2.

```{r setup}
library(tidyverse)
library(readxl)
```

# Question 1
### Part 1: Read and clean the Mr. Trash Wheel sheet
(1) Specify the sheet in the Excel file and to omit non-data entries
(2) Use reasonable variable names
(3) Omit rows that do not include dumpster-specific data
(4) Round the number of sports balls to the nearest integer and converts the result to an integer variable
```{r readnclean_trashwheel}
trashwheel_df = 
	read_xlsx(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

### Part 2: Read and clean precipitation data for 2017 and 2018
(1) For each, omit rows without precipitation data and add a variable year.
```{r readnclean_precipitation}
precip_2017 = 
	read_xlsx(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)

precip_2018 = 
	read_xlsx(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
```

(2) Combine precipitation datasets and convert month to a character variable.
```{r combine_precipitation}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(precip_2017, precip_2018)

precip_df =
	left_join(precip_df, month_df, by = "month")

precip_df %>% select(year, month = month_name,total)
```

### Part 3: Write a paragraph about these data
Mr. Trash Wheel dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`

Additional data sheets include month precipitation data, in which:

* There are a total of `r nrow(precip_df)` rows in our final dataset that combine precipitation data for 2017 and 2018. 
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


# Question 2
### Part 1: Read and clean NYC Transit data
Retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
```{r readnclean_transit}
transit_df = 
	read_csv(
		"./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
	janitor::clean_names() %>%
  select(line:station_longitude, route1:route11, entry, vending, entrance_type, ada)
```

### Part 2: Convert the entry variable from character to a logical variable
```{r logical_entry}
transit_df = 
  transit_df %>%
  mutate(entry = if_else(entry == "YES", TRUE, FALSE))
```

### Part 3: Write a short paragraph about this dataset
NYC Transit data contains information related to each entrance and exit for each subway station in NYC. Data also include information on lines, stations and routes. So far, my data cleaning steps include cleaning the variable names, selecting required variables and converting _entry_ from character to a logical variable. There are a total of `r nrow(transit_df)` rows x `r ncol(transit_df)` columns in our final dataset. Data are not tidy yet, because route served is spread across 11 columns.

### Part 4: Answer the following questions
(1) How many distinct stations are there?\
There are `r transit_df %>% distinct(line, station_name) %>% nrow()` distinct stations.

(2) How many stations are ADA compliant?\
There are `r filter(transit_df, ada == TRUE) %>% distinct(line, station_name) %>% nrow()` stations that are ADA compliant.

(3) What proportion of station entrances / exits without vending allow entrance?\
`r transit_df %>% filter(vending == "NO", entry == TRUE) %>% nrow() / transit_df %>% filter(vending == "NO") %>% nrow()` of station entrances / exits without vending allow entrance.

### Part 5: Reformat data so that route number and route name are distinct variables
```{r distinct_route}
transit_tidy = 
	transit_df %>%
  mutate_at(vars(route1:route11), as.character) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) %>%
  drop_na(route_name)
```
(1) How many distinct stations serve the A train?\
`r transit_tidy %>% distinct(line, station_name,route_name) %>% filter(route_name == "A") %>% nrow()` distinct stations serve the A train.

(2) Of the stations that serve the A train, how many are ADA compliant?\
Of the stations that serve the A train, `r transit_tidy %>% distinct(line, station_name, route_name, ada) %>% filter(route_name == "A", ada == TRUE) %>% nrow()` are ADA compliant.


# Question 3
### Part 1: Read and clean data in pols-month.csv
(1) Break up the variable mon into integer variables year, month, and day
(2) Replace month number with month name
(3) Create a president variable taking values gop and dem, and remove prez_dem and prez_gop
(4) Remove the day variable
```{r readnclean_pols}
pols_df = 
	read_csv(
		"./data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"))

month_df = 
	tibble(
		month = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12"),
		month_name = month.name
	)

pols_df =
	left_join(pols_df, month_df, by = "month")  %>%
  mutate(month = month_name) %>%
  select(-month_name) %>%
  mutate(president = case_when(prez_gop == 1 ~ "gop", prez_dem == 1 ~ "dem")) %>%
  select(-c(prez_gop, prez_dem, day)) %>%
  relocate(year, month, president)
```

### Part 2: Read and clean data in snp.csv
For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.
```{r readnclean_snp}
snp_df = 
	read_csv(
		"./data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into = c("month", "day", "year")) %>%
  select(-day) %>% 
  relocate(year, month)
  

month_df = 
	tibble(
		month = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
		month_name = month.name
	)

snp_df =
	left_join(snp_df, month_df, by = "month")  %>%
  mutate(month = month_name) %>%
  select(-month_name)
```

### Part 3: Read and clean data in unemployment.csv
(1) Switching from “wide” to “long” format
(2) Ensuring that key variables have the same name
(3) Ensuring that key variables take the same values.
```{r readnclean_unemp}
unemp_df = 
	read_csv(
		"./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_pct"
  ) %>%
  rename(year = Year) %>% 
  mutate(year = as.character(year))
  
month_df = 
	tibble(
		month = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
		month_name = month.name
	)

unemp_df =
	left_join(unemp_df, month_df, by = "month")  %>%
  mutate(month = month_name) %>%
  select(-month_name)
```

### Part 4: Join the datasets by merging snp into pols, and merging unemployment into the result
```{r merge}
pols_snp_df =
	left_join(pols_df, snp_df, by = c("year", "month"))

pols_snp_unemp_df =
  left_join(pols_snp_df, unemp_df, by = c("year", "month"))
```

### Part 5: Write a short paragraph about these datasets
My result dataset was joined by three datasets: pols-month, snp, and unemployment. Before merging these datasets by _year_ and _month_, I made sure that the formats and data types of _year_ and _month_ variables were consistent across all three datasets. pols-month dataset contains the number of national politicians who are democratic or republican at given time. The indicators of whether the president was democratic or republican on the associated date were recoded by a new variable called _president_. snp dataset contains _close_ variable, which indicates the closing values of the S&P stock index on the associated date. umemployment dataset contains _unemployment_pct_ variable, which indicates the percentage of unemployment at given time. There are a total of `r nrow(pols_snp_unemp_df)` rows x `r ncol(pols_snp_unemp_df)` columns in our final dataset.The range of _year_ is [`r range(pull(pols_snp_unemp_df, year))`]. 

